Large and Infinitary Quotient Inductive-Inductive Types

### 2020-01-02 draft/paper.tex
paper

\newcommand{\ul}[1]{\underline{#1}}
\newcommand{\ulGamma}{\ul{\Gamma}}
\newcommand{\ulDelta}{\ul{\Delta}}
\newcommand{\ulgamma}{\ul{\gamma}}
\newcommand{\uldelta}{\ul{\delta}}
\newcommand{\ulsigma}{\ul{\sigma}}
\newcommand{\ulepsilon}{\ul{\epsilon}}
\newcommand{\ult}{\ul{t}}
\newcommand{\ulu}{\ul{u}}
\newcommand{\ulA}{\ul{A}}
\newcommand{\ulB}{\ul{B}}

@@

\begin{definition}[Signature for ToS] For each level $i$,
we define $\ToSSig_i : \Sig_{1+i}$, as the signature for the theory of
signatures with external sets in $\Set_i$. This is a large and infinitary QIIT
signature, as we have $\Pie$ and $\Piinf$ abstracting over $A : \Set_i$ and
branching with $A \ra \Ty\,\Gamma$ and $A \ra \Tm\,\Gamma\,\U$ respectively. We
present an excerpt from $\ToSSig_i$ below.
\begin{alignat*}{3}
  & \emptycon && \ext (Con : \U)\\
  &           && \ext (Sub : Con\arri Con \arri \U)\\
  &           && \ext (Ty  : Con\arri\U)\\
  &           && \ext (Tm  : \Pii(\Gamma : Con)(\Ty\,\Gamma \arri \U))\\
  &           && ...\\
  &           && \ext (\Piinf : \Pii(\Gamma : Con)\\
  &           && \hspace{4.2em}(\Pie \Set_i (\lambda\,A.\,(A \arrinf \Ty\,\Gamma)\arri\El(\Ty\,\Gamma))))\\
  &           && ...
\end{alignat*}

Now, for each $j$, the interpretation of $\ToSSig_i$ in $\bM_{1+i,j}$ yields an
flCwF $\bGamma$ such that $\Con_{\bGamma} = \ToS_{i,j}$. In short, we can
recover ToS algebras from $\ToSSig$. This follows by computation of the
interpretation and the fact that $\ToSSig$ is the internal representation of
$\ToS$. Also, since we have an flCwF of $\ToS_{i,j}$-algebras, we can use
Definition \ref{def:induction} for the notion of $\ToS$-induction.

@@

%% \begin{theorem}  For each $j$, the interpretation of $\ToSSig_i$ in $\bM_{1+i,j}$
%% yields an flCwF $\bGamma$ such that $\Con_{\bGamma} = \ToS_{i,j}$. In short, we
%% can recover ToS algebras from $\ToSSig$.
%% \end{theorem}
%% \begin{proof}
%% \end{proof}

%% Overall, this is just rephrasing Definition \ref{def:tos}
%% internally to an arbitrary $M : \ToS_{1+i,j}$; recall from Definition
%% \ref{def:signature} that $\Sig_{1+i} = (j : \mathbb{N})(M : \ToS_{1+i,j})\ra
%% p\Con_M$. For example, the underlying sets are specified as

>>>>

\newcommand{\ul}[1]{\underline{#1}}
\newcommand{\ulGamma}{\ul{\Gamma}}
\newcommand{\ulDelta}{\ul{\Delta}}
\newcommand{\ulgamma}{\ul{\gamma}}
\newcommand{\uldelta}{\ul{\delta}}
\newcommand{\ulsigma}{\ul{\sigma}}
\newcommand{\ulepsilon}{\ul{\epsilon}}
\newcommand{\ult}{\ul{t}}
\newcommand{\ulu}{\ul{u}}
\newcommand{\ulA}{\ul{A}}
\newcommand{\ulB}{\ul{B}}
\newcommand{\tos}{\mathsf{tos}}

@@

\begin{definition}[Signature for ToS] For each level $i$,
we define $\ToSSig_i : \Sig_{1+i}$, as the signature for the theory of
signatures with external sets in $\Set_i$. This is a large and infinitary QIIT
signature, as we have $\Pie$ and $\Piinf$ abstracting over $A : \Set_i$ and
branching with $A \ra \Ty\,\Gamma$ and $A \ra \Tm\,\Gamma\,\U$ respectively. We
present an excerpt from $\ToSSig_i$ below.
\begin{alignat*}{3}
  & \emptycon && \ext (Con : \U)\\
  &           && \ext (Sub : Con\arri Con \arri \U)\\
  &           && \ext (Ty  : Con\arri\U)\\
  &           && \ext (Tm  : \Pii(\Gamma : Con)(\Ty\,\Gamma \arri \U))\\
  &           && ...\\
  &           && \ext (\Piinf : \Pii(\Gamma : Con)\\
  &           && \hspace{4.2em}(\Pie \Set_i (\lambda\,A.\,(A \arrinf \Ty\,\Gamma)\arri\El(\Ty\,\Gamma))))\\
  &           && ...
\end{alignat*}
\end{definition}

Now, for each $j$, the interpretation of $\ToSSig_i$ in $\bM_{1+i,j}$ yields an
flCwF $\bGamma$ such that $\Con_{\bGamma} = \ToS_{i,j}$. In short, we can
recover ToS algebras from the semantics of $\ToSSig$. This follows by
computation of the interpretation and the fact that $\ToSSig$ is precisely the
internal representation of $\ToS$. Also, since we have an flCwF of
$\ToS_{i,j}$-algebras, we can use Definition \ref{def:induction} for the notion
of $\ToS$-induction.

@@

In this section we construct QIITs from initial ToS-algebras. From now on, we assume
the existence of such algebras.

\todo{assumption in environment}

\textbf{For each $i$, we assume that $\tos_i : \ToS_{i, 1+i}$ exists}. The reason for the
$1 + i$ level of $\ToS$ sorts is that $\tos_i$ may contain $A : \Set_i$ types in
$\Ty$, and since we intend $\tos_i$ to be a freely generated algebra, the sorts
must be large enough to avoid size paradoxes.

\begin{lemma}[Cumulativity of $\ToS$] If $i \leq i'$ and $j \leq j'$,
  then $\ToS_{i,j} \leq \ToS_{i', j'}$.
\end{lemma}
\begin{proof} $\ToS_{i,j}$ is uniformly defined in $i$, $j$, hence cumulative subtyping extends
  through dependent pairs and functions to $\ToS$.
\end{proof}

%% \todo{assumption as environment}

%% \emph{Assumption.}
